Spreading of a Facebook Application

### 2011-04-15 paper.tex
checked abstract to results - metrics

\subsection{Data evaluation}
\label{sec:data-evaluation}
Data evaluation is performed by using the MySQL database, and
PHP scripts. In addition, the dot
language\footnote{http://www.graphviz.org/Documentation.php} is employed
to create graphs.

The evaluation scripts are maintained by \textit{main.php}. A user
class represents Facebook's social graph internally. Its properties
are user's Facebook ID, registration date, average score, a list of
friends, and a list of IDs to whom the user may have spread the
application. Friendships are represented by a list of references to corresponding
user objects, and names of users' can be made anonymous if needed.

A user is supposed to have spread the app if he/she has
a friend who used the application and who belongs to a \textit{new}
network. Such persons are found by recursively checking if an user of
the application has friends, who also used the application, and whose friends
also used the app, but are not in his/her list of friends, and also not in the list of the person
who originally spread the application. In this case, the corresponding
friend will be marked as a person who spread the application, and the
recursion starts by letting the friend become the analyzed user.
Since it is not clear, why a person started to use an application, the
script also allows for identification of users who have friends 
who also participated in the application.

User names are made anonymous by substituting their names with a
random name selected from a list of popular baby names
from 2009, taken from \url{http://www.ssa.gov/cgi-bin/popularnames.cgi}.   
For each gender, hundred names are stored in an array, and if the
gender is set, a name of same gender is chosen. Otherwise, the name is
taken randomly. The name of the person who spread the application
originally is always set to \textit{Alexandra}.

Output options are a \ac{csv}-file containing metrics calculated from
the internal social graph as well as various graphs in dot language, created
from the internal social graph, too.
% Properties of
% user objects are in the style of the \textit{user}-table's properties, and a user's
% . The whole set of users is maintained by
% \textit{main.php}, where they are stored in an array. Class \textit{Graph}
% allows for output of various graphs in dot-language created from the
% internal social graph.

\section{Results}
\label{sec:results}
A total of n=52 users participated in the quiz. 6311 users have been
stored in the user table. Thus, on average each user brought on
average 121.4 \textit{new} people, who could read about the application
in their news stream.

\subsection{Characteristics}
\label{sec:characteristics}

The characteristics discussed in this section consider data of users
that are still registered (n=37).

\subsubsection{Age and gender}
\label{sec:age-gender}
$59 \%$ of the registered user are female, $41 \%$ of the participants
are male. Table \ref{tab:agegender} shows classification of the users
by age and gender. It can be seen, that most users ( 70 \%) were at
the age of 25-34. Only few ($8.4 \%$) were older, and only 21.4 \%
were younger.

\begin{center}
\begin{table}
\label{tab:agegender}
\begin{tabular}{l | c c c c}
Gender\textbackslash Age  & 13-17 & 18-24 & 25-34 & 45-54 \\
\hline
female  & 2.7 \% & 16 \% & 35 \% & 5.4 \% \\
male & 0 \% & 2.7 \% & 35 \% & 2.7 \% 
\end{tabular}
\caption{Table shows classification of users by age and gender.}
\end{table}
\end{center}

\subsubsection{Demography}
\label{sec:demographie}
Table \ref{tab:countries} details the origin of the users and table
\ref{tab:languages} the languages in which the participants use Facebook.
As it seems obvious for a quiz in German, most visitors came from Germany.

\begin{center}
\begin{table}
\label{tab:countries}
\begin{tabular}{l | c }
Country  & Visitors \\
\hline
Germany  & 32 \\
United Kingdom & 2 \\
Israel & 1 \\
United States of America & 1 \\
Austria & 1 \\
\end{tabular}
\caption{Table shows where the users are from.}
\end{table}
\end{center}

\begin{center}
\begin{table}
\label{tab:languages}
\begin{tabular}{l | c }
Languages   & Visitors \\
\hline
German  & 25 \\
English (UK) & 7 \\
English (US) & 4 \\
Greek  & 1 \\
\end{tabular}
\caption{Languages of the participants.}
\end{table}
\end{center}

@@

\newpage
\section{Addendum}
\label{sec:addendum}

@@

@@

 \begin{table}
 \label{tab:spread}
 \begin{tabular}{ c c c c}
Friends & Score & Friends who also participated & Probably spread appl. to friends\\
\hline
97 & 3.33 & 1 &-\\
322 & 1.5 & 4 & -\\
63 & 4 & 1 & 1\\
102 & 0 & 2 & -\\
91 & 1 & 3 &1\\
128 & 4 & 2 & -\\
151 & 3 & 4 &3\\
135 & 3.64 & 3 &2\\
254 & 4 & 2 & 2\\
220 & 3 & 4 & -\\
127 & 3 & 2 & -\\
186 & 4 & 3 & -\\
42 & 3 & 4 & - \\
445 & 3 & 4 & - \\
153 & 3 & 2 &1\\
20 & 3 & 1 &1\\
11 & 4.23 & 8 &2\\
130 & 3 & 1 & -\\
67 & 2 & 1 & - \\
207 & 2 & 2 & -\\
74 & 4.5 & 3 & -\\
47 & 4.42 & 3& -\\
157 & 3 & 2 & -\\
136 & 3 & 2 & -\\
78 & 4 & 1 & -\\
167 & 2 &1 & -
\end{tabular}
\caption{Table shows how many friends and which score user had who may have spread the
  application, and how many of their friends started to use the
  application. If a user was a connector to another network, it is
  calculated to how many new persons the application (appl.) was
  probably spread.}
\end{table}

 \begin{center}
 \begin{table}
 \label{tab:nospread}
 \begin{tabular}{ c c c}
Friends & Score  \\
\hline
90 & 1.29 \\
332 & 3 \\
233 & 4 \\
79 & 3 \\
53 & 1 \\
89 & 2 \\
72 & 4 \\
76 & 1 \\
264 & 1 \\
35 & 3 \\
121 & 4 \\
289 & 1 \\
15 & 4 \\
49 & 4 \\
0 & 4.75 \\
80 & 4 \\
60 & 0 \\
13 & 2.5 \\
229 & 0 \\
53 & 0 \\
10 & 3 
\end{tabular}
\caption{Table shows how many friends and which score user had who did
  not spread the application.}
\end{table}
\end{center}

>>>>

\subsection{Data evaluation}
\label{sec:data-evaluation}
For data evaluation, a tool coded in PHP exists. It allows for output
of metrics as \ac{csv} and graphs in \ac{dot}
language\footnote{http://www.graphviz.org/Documentation.php}.
The data evaluated by the tool are taken from
 database tables the quiz also makes use of.

As mentioned in previous sections, the tool maintains an extract of
Facebook's social graph internally. This graph contains users' basic
information, friendships, etc. The internal graph is
extended by information from the quiz, such as if the user has
installed the application, and which score a user had. The internal graph
consists of instances of class \textit{user}, which have references
among each other representing friendships. The whole set of users is
maintained by \textit{main.php}, which contains a list of all the user
instances.

Before the graph is created, user names can be made anonymous and the
results can be stored in the \textit{user} table to provide
reusability, too. User names are made anonymous by random selection of
a name from a list of popular baby names from 2009. The list of names was
taken from \url{http://www.ssa.gov/cgi-bin/popularnames.cgi}.   
For each gender, hundred names are stored in an array, and if the
gender is set, a name of same gender is chosen. Otherwise, the name is
taken randomly. The name of the person who spread the application
originally is always set to \textit{Alexandra}.

The tool can make a prognoses, which users spread the quiz. To
identify such users, the following idea is employed:
A user is supposed to have spread the app if he/she has
friends who also used the application and who belong to a \textit{new}
social network. Figure \ref{fig:newNW} gives an example.
The implementation is as follows: People who are supposed to have
spread the quiz are found by recursively checking if an user of 
the application has friends, who also used the application, and whose friends
also used the app, but are not in his/her list of friends, and also not in the list of the person
who originally spread the application. In this case, the corresponding
friend will be marked as a person who spread the application, and the
recursion starts by letting the friend become the analyzed user.
Since it is not clear, why a person started to use an application, the
script also allows for identification of users who have friends 
who also participated in the application.

To let the tool output the graphs in \ac{dot} language, a parameter
has to be given to the tool via HTTP-Get method. The parameters are:
\verb|spreading|, \verb|tree|, \verb|friendsFilled| and
\verb|friendsScored|. If type \verb|spreading| is selected, only users
who participated in the quiz are shown. Users who are supposed to have
spread the application are red colored. Next to each user, the number
of friends, and the score the user yielded in the quiz are termed. If
type \verb|tree| is 
selected, all the users are shown, and nodes of users who participated
in the quiz are red colored. If type \verb|friendsFilled| is selected,
only users who participated in the quiz are considered and the
background color is set corresponding to the number of friends. A user
with the highest number of friends has background color black, the
person with the lowest number of friends has background color
white. If type \verb|friendsScored| is selected, the graph is created
analogous to type \verb|friendsFilled|, but the background color is
computed with respect to the score user had in the quiz averagely.

To let the tool output \ac{csv}, the \verb|csv| parameter has to be
set to \verb|appUser| or to \verb|noAppUser|. 
If \verb|appUser| is selected, for each user the following metrics are
calculated: The number of friends, the number of friends who
participated in the application, and to how many people a person
probably spread the application.  If \verb|noAppUser| is selected, it
is only computed how many friends these user had who participated in
the quiz. It is returned, how many user had one friend, two friends,
and so on.

Besides the tool, some data are directly retrieved using database
queries. These data include the number of people in the user table, and
the number of people who participated in the application.

% Output options are a \ac{csv}-file containing metrics calculated from
% the internal social graph as well as various graphs in dot language, created
% from the internal social graph, too.
% Properties of
% user objects are in the style of the \textit{user}-table's properties, and a user's
% . The whole set of users is maintained by
% \textit{main.php}, where they are stored in an array. Class \textit{Graph}
% allows for output of various graphs in dot-language created from the
% internal social graph.

\section{Results}
\label{sec:results}
A total of n=52 users participated in the quiz and 6311 users have been
stored in the user table. Data were collected from 15th of February to
15th of April. This section presents, which
characteristics these users have based on Facebook's insights and on
metrics computed with the tool discussed in section
\ref{sec:data-evaluation}. Then, the graphs detailing the spreading of
the application are shown.

\subsection{Characteristics Of the Participants}
\label{sec:characteristics}
In this section, characteristics about the users who participated in
the quiz are given. At first, characteristics retrieved from
Facebook's
\textit{Insights}\footnote{http://www.facebook.com/help/?search=insights}
are described, than, metrics collected from database and by the tool are given.

\subsubsection{Facebook's Insights}
Facebook Insights `provides Facebook Page owners and Facebook Platform
developers with metrics around their content.'\cite{facebookInsights}
These metrics are free and can be found in the \textit{My Apps}
section on the Facebook Developer Platform.
Facebook Insights considers for its metrics life time installation of an application
and events happened in a month. In case of the analyzed application,
Facebook had 50 Lifetime installed users for the quiz on 15th of April, 2011.

\paragraph{Age and gender}
\label{sec:age-gender}
$60 \%$ of the registered user are female, $40 \%$ of the participants
are male. Table \ref{tab:agegender} shows classification of the users
by age and gender. It can be obtained, that most users ( 68 \%) were at
the age of 25-34. Few participants ($8 \%$) were older, and about a
quarter (24 \%) were younger than 25.

\begin{center}
\begin{table}
\label{tab:agegender}
\begin{tabular}{l | c c c c c}
Gender\textbackslash Age  & 13-17 & 18-24 & 25-34 & 45-54 & 55+\\
\hline
female  & 2 \% & 18 \% & 36 \% & 4 \% & 0 \%\\
male & 0 \% & 4 \% & 32 \% & 2 \%  & 2 \\
\end{tabular}
\caption{Table shows classification of users by age and gender.}
\end{table}
\end{center}

\paragraph{Demography}
\label{sec:demographie}
Table \ref{tab:countries} details the origin of the users and table
\ref{tab:languages} the languages in which the participants chose
Facebook to occur.

\begin{center}
\begin{table}
\label{tab:countries}
\begin{tabular}{l | c }
Country  & Visitors \\
\hline
Germany  & 44 \\
United Kingdom & 2 \\
Israel & 1 \\
United States of America & 1 \\
Austria & 1 \\
\end{tabular}
\caption{Table shows where the users are from.}
\end{table}
\end{center}

\begin{center}
\begin{table}
\label{tab:languages}
\begin{tabular}{l | c }
Languages   & Visitors \\
\hline
German  & 34 \\
English (US) & 10 \\
English (UK) & 4 \\
Greek  & 1 \\
Italian & 1 \\
\end{tabular}
\caption{Languages of the participants.}
\end{table}
\end{center}

\paragraph{Requests}
The quiz allows for invitation of friends to participation by sending
requests. In the time of 18th of March to 8th of April, 4 requests
were sent, and one was accepted. In the month before, no friend
requests were accepted or ignored.

\subsubsection{Metrics collected by the tool}
As detailed in section \ref{sec:data-evaluation}, the tool allows for
output of metrics as \ac{csv}. Metrics about users who participated in
the quiz are given in table
\ref{tab:spread}. Table \ref{tab:nouser}
describes user who did not participate. Table \ref{tab:scorespread}
and \ref{tab:nospread} sum up metrics from table \ref{tab:spread}.

Five users participated in the application, but

@@

\newpage
\section{Tables and Figures}
\label{sec:addendum}

@@

\begin{figure}
  \includegraphics[height=3in]{../img/newNW.png}
 \caption{This figure emphasizes users who connect to new social networks with
   red border.}
\label{fig:newNW}
\end{figure}

@@

 \begin{table}
 \label{tab:spread}
 \begin{tabular}{ c c c c c}
Anon. Name&Friends&App-friends&Spread App to&Average Score	\\
Alexandra & 133& 34& 34 & 4.4625 \\
Evan & 97& 1& 0 & 3.33333 \\
Justin & 90& 2& 0 & 1.28571 \\
Tristan & 332& 2& 0 & 3 \\
Thomas & 396& 3& 0 & 4 \\
Jose & 233& 1& 0 & 4 \\
David & 322& 6& 0 & 1.5 \\
Gianna & 63& 3& 1 & 4 \\
Makayla & 79& 1& 0 & 3 \\
Connor & 102& 3& 0 & 0 \\
Ethan & 91& 4& 1 & 1 \\
Isabelle & 53& 1& 0 & 1 \\
Gavin & 89& 1& 0 & 2 \\
Caleb & 128& 3& 0 & 4 \\
Sarah & 151& 5& 3 & 3 \\
Emma & 72& 1& 0 & 4 \\
Gabrielle & 135& 4& 2 & 3.63636 \\
David & 254& 3& 2 & 4 \\
Emily & 220& 5& 0 & 3 \\
Victoria & 63& 2& 0 & 2 \\
Brian & 126& 2& 0 & 4 \\
Jose & 127& 3& 0 & 3 \\
Thomas & 186& 4& 0 & 4 \\
Evelyn & 42& 5& 0 & 3 \\
Paige & 76& 2& 0 & 1 \\
Hannah & 264& 2& 0 & 1 \\
Charlotte & 445& 5& 0 & 3 \\
Elizabeth & 35& 1& 0 & 3 \\
Carson & 153& 3& 1 & 3 \\
Carson & 20& 2& 1 & 3 \\
Nathan & 121& 1& 0 & 4 \\
Eli & 11& 9& 2 & 4.23077 \\
Isaac & 290& 1& 0 & 1 \\
Alexandra & 130& 1& 0 & 3 \\
Gabriel & 67& 1& 0 & 2 \\
Paige & 16& 1& 0 & 4 \\
David & 50& 1& 0 & 4 \\
Julian & 207& 2& 0 & 2 \\
Carlos & 3& 3& 0 & 4.75 \\
Evan & 74& 3& 0 & 4.5 \\
Sebastian & 47& 3& 0 & 4.41667 \\
Xavier & 80& 0& 0 & 4 \\
Evan & 157& 2& 0 & 3 \\
Brody & 136& 2& 0 & 3 \\
Luis & 60& 1& 0 & \\
Allison & 13& 0& 0 & 2.5 \\
Tyler & 229& 0& 0 & \\
Emily & 53& 0& 0 & \\
Khloe & 78& 1& 0 & 4 \\
Katelyn & 167& 1& 0 & 2 \\
Kaitlyn & 10& 1& 0 & 3 \\
Eva & 272& 0& 0 & 4
\end{tabular}
\caption{Table details metrics about users who participated in the
  application. Friends detail the number of friends a user had,
  App-friends is the number of friends who also participated in the
  quiz, Spread App To details to how many people a user probably
  spread the application, and Average Score details to the score the
  user averagely had in the quiz. The names were made anonymous.}
\end{table}

 \begin{center}
 \begin{table}
 \label{tab:nospread}
 \begin{tabular}{ c c}
Number of app friends & Occurrences \\
\hline
0 app friends & 5 \\
1 app friend & 17 \\
2 app friends & 10 \\
3 app friends & 10 \\
4 app friends & 3 \\
5 app friends & 3\\
6 app friends & 2\\
9 app friends & 1\\
34 app friends & 1 \\
\end{tabular}
\caption{Table sums up metrics from table \ref{tab: spread}. It
  details how many users had 0 friends who also participated in the
  application, one friend, and so further.}
\end{table}
\end{center}

\begin{table}
\label{tab:scorespread}
\begin{tabular}{c c}
Spread to & Average score \\
1&	2.75 \\
2&	3.96 \\
3&	3 \\
\end{tabular}
\caption{Table details the average score of people who spread the
  application to one, two, or three users.}
\end{table}
